---
title: "Visualization with ggplot 1"
author: "Esther Kyeremah"
date: "2023-10-22"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


##Scatterplts!!
Create my first scatterplot

```{r}
weather_df |>
ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Save and edit a plot object
```{r}
weather_plot=
weather_df |>
ggplot(aes(x = tmin, y = tmax)) 

weather_plot + geom_point()
```

##Advanced scatterplot..
Start with th 
```{r}
weather_df |>
ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```

What aout the aes
```{r}
weather_df |>
ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) +
  geom_smooth()
```

Let's facet some things!!
```{r}
weather_df |>
ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

```{r}
weather_df |>
ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)
```

Define alpha(here it means size/transparency) level
```{r}
weather_df |>
ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .2, size = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

```
 Lets try a new pkot with date
 
```{r}
weather_df |>
ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp),alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```
 
 How many geoms have to exist?
 you can have whatever geoms you want but 
 below may not be useful
 
```{r}
weather_df |>
ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```
 
You can use a neat geom! this is overkill for a scatterplot with 1000 observations but for very large datasets this helps tbecause changing your alpha in those cases sometimes doesnt help to see


```{r}
weather_df |>
ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_hex()
```

adding on geom density
```{r}
weather_df |>
ggplot(aes(x = tmin, y = tmax)) + 
geom_density2d() +
geom_point(alpha = .3)
```

##univariate plots

```{r}
weather_df |>
ggplot(aes(x = tmin)) +
  geom_histogram()
```

when you want to add color

```{r}
weather_df |>
ggplot(aes(x = tmin, color = name)) +
  geom_histogram()
```

DOESNT LOOK TOO GOOD COS THEYRE JUST STACKED ON EACH OTHER AND THE COLOR IS ON THE OUTSIDE OF THE BAR

```{r}
weather_df |>
ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```

EVEN THOUGH THEYRE NOT REALLY STACKED THE DODGE DOESNT STILL MAKE IT TOO REPRESNTABLE

so..

```{r}
weather_df |>
ggplot(aes(x = tmin, fill = name)) +
geom_histogram() +
  facet_grid(. ~ name)
```

Let's try a new geometry! 
this one makes you see more or less bumps in the histogram
```{r}
weather_df |>
ggplot(aes(x = tmin, fill = name)) +
geom_density(alpha = .3, adjust = .5)
```


Lets try box plots

```{r}
weather_df |>
ggplot(aes(x = name, y = tmin)) +
geom_boxplot()
```


Trendy plots
```{r}
weather_df |>
ggplot(aes(x = name, y = tmin, fill = name)) +
geom_violin( alpha = .5) +
stat_summary(fun = "median")
```

Ridge plots__ the most popular plot of 2017
```{r}
weather_df |>
ggplot(aes(x = tmax, y = name, fill = name)) +
geom_density_ridges()
```

# Save and embed

Lets save a scatterplot

```{r}
weather_save = 
weather_df |>
ggplot(aes(x = tmin, y = tmax, color = name)) +
geom_point(alpha = .5)

ggsave("./results/weather_save.pdf", weather_save, width = 8, height = 5)

```

What about embedding
Embed at different size

```{r, fig.width=6, fig.height=3 }
weather_save
```


